---
title: "ATAC-Seq-Data-Analysis-QC on all cells Step 2: Visualisation of LogNormalization"
date: "`r Sys.Date()`"
author: Xenia Sirodzha
output:
  html_document: 
    toc: yes
    toc_depth: 4
    toc_float: yes
    df_print: paged
---

# Libraries

```{r}
library(tidyverse) 
library(binom)
library(lubridate)
library(broom)
library(scales) #for scale transformations
library(cowplot) #arranging plots 
library(ggpubr)
library(ggridges)
library(ggthemes) #stylising plots
library(ggplot2)
library(ggsci) #color palettes
library(readxl) #for loading excel files
library(Seurat) #for rna-seq data, atac-seq data
library(Signac) #for atac seq data
library(sparsevctrs)
```

# 1 Data Loading

```{r}
readRDS('../data/Cell-type_Peak_Matrix.rds') -> zfish_snATAC_seq_pk_mtrx
readRDS("../data/psd.bulk.zfish_atac.lg1x.rds") -> psd.bulk.zfish_atac.lg1x
psd.bulk.zfish_atac.lg1x[1:5,1:5]
```


# 2 Data Transformation for Plotting
After normalisation, I prepared the data to plot with tidyverse. First, I converted the matrix to a data frame, then I calculated the values which will be plotted as well(sum of pk_counts, coefficient of variance (cv))

```{r}
as.data.frame(psd.bulk.zfish_atac.lg1x)%>% 
  pivot_longer(cols=!contains("rownames"), ,values_to = "pk_counts", names_to = "celltype") %>% 
  mutate(pk_count_sum= sum(pk_counts)) %>% 
  mutate(cv=sd(pk_counts) / mean(pk_counts)) %>% 
  group_by(celltype) %>% 
  mutate(pk_count_sum_celltype= sum(pk_counts)) %>% 
  mutate(cv_celltype=sd(pk_counts) / mean(pk_counts)) %>% 
  mutate(all="all celltypes")-> df_psd.bulk.zfish_atac.lg1x

df_psd.bulk.zfish_atac.lg1x
```

```{r eval=FALSE}
df_pk_mtrx= as_tibble(zfish_snATAC_seq_pk_mtrx, rownames="rownames")
df_pk_mtrx %>% pivot_longer(cols=!contains("rownames"), values_to = "pk_counts", names_to = "id") -> df_pk_mtrx.l
df_pk_mtrx.l
```



# 3 Plotting

## 3.1 all cells counts as a histogram 

This is not possible with all cells.

```{r eval=FALSE}
df_pk_mtrx.l %>% 
  ggplot(aes(x=pk_counts))+geom_histogram(binwidth = 0.7, color="gray15", alpha=0.8)+
  theme_bw()+ theme(text = element_text(size = 9, family="sans", color="black"), #stylisation
        axis.text = element_text(size=9, color="black"),
        axis.title=element_text(size=9, face = "bold"),
        plot.subtitle=element_text(size=10.5),
        plot.title=element_text(size=11, face = "bold"), 
        legend.position ="none")
```




## 3.2 Distribution After pseudo-bulking by celltype

```{r}
df_psd.bulk.zfish_atac.lg1x %>% 
  ungroup() %>% 
 ggplot(aes(x=pk_counts))+geom_histogram(binwidth = 0.2, color="gray15", alpha=0.8)+
  theme_bw()+ theme(text = element_text(size = 9, family="sans", color="black"), #stylisation
        axis.text = element_text(size=9, color="black"),
        axis.title=element_text(size=9, face = "bold"),
        plot.subtitle=element_text(size=10.5),
        plot.title=element_text(size=11, face = "bold"), 
        legend.position ="none") 
```



```{r}
df_psd.bulk.zfish_atac.lg1x %>% 
 ggplot(aes(x=pk_counts))+geom_histogram(binwidth = 0.2, color="gray15", alpha=0.8)+
 scale_y_continuous(breaks = scales::breaks_pretty(n=10))+
  coord_fixed(ratio = 0.00001)+
  theme_bw()+ theme(text = element_text(size = 9, family="sans", color="black"), #stylisation
        axis.text = element_text(size=9, color="black"),
        axis.title=element_text(size=9, face = "bold"),
        plot.subtitle=element_text(size=10.5),
        plot.title=element_text(size=11, face = "bold"), 
        legend.position ="none") ->dis.counts.1

 ggplot(df_psd.bulk.zfish_atac.lg1x, aes(x=pk_counts, fill = celltype))+
  geom_histogram(binwidth = 0.2, color="gray15", alpha=0.8)+
  scale_y_continuous(breaks = scales::breaks_pretty(n=10))+
  facet_wrap(facets = "celltype")+
  theme_bw()+ theme(text = element_text(size = 9, family="sans", color="black"), #stylisation
        axis.text = element_text(size=9, color="black"),
        axis.title=element_text(size=9, face = "bold"),
        plot.subtitle=element_text(size=10.5),
        plot.title=element_text(size=11, face = "bold"), 
        legend.position ="none") ->dis.counts.cell.1

dis.counts.cell.1
```

```{r}
dis.counts.1|dis.counts.cell.1
```



## 3.3 Sum of all peaks

```{r}
unique(df_psd.bulk.zfish_atac.lg1x$pk_count_sum)
```


```{r}

df_psd.bulk.zfish_atac.lg1x %>% 
  ungroup() %>% 
  distinct(all, pk_count_sum) %>% 
  ggplot(aes(x=all, y=pk_count_sum))+
  geom_col(alpha=0.8, color="gray10")+
  coord_fixed(ratio = 0.0000005)+
  theme_bw()+ theme(text = element_text(size = 9, family="sans", color="black"), #stylisation
        axis.text = element_text(size=9, color="black"),
        axis.title=element_text(size=9, face = "bold"),
        plot.subtitle=element_text(size=10.5),
        plot.title=element_text(size=11, face = "bold"), 
        legend.position ="none") -> sum.peaks.all.1


df_psd.bulk.zfish_atac.lg1x %>% 
  group_by(celltype) %>% 
  distinct(pk_count_sum_celltype, pk_count_sum) %>% 
  ggplot(aes(x=celltype, y=pk_count_sum_celltype, fill=celltype))+
  geom_col(alpha=0.8, color="gray10")+
  theme_bw()+ theme(text = element_text(size = 9, family="sans", color="black"), #stylisation
        axis.text = element_text(size=9, color="black"),
        axis.title=element_text(size=9, face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        plot.subtitle=element_text(size=10.5),
        plot.title=element_text(size=11, face = "bold"), 
        legend.position ="none") -> sum.peaks.celltype.1

sum.peaks.all.1|sum.peaks.celltype.1
```


## 3.4 Coefficient of Variation 


```{r}
df_psd.bulk.zfish_atac.lg1x %>% 
  ungroup() %>% 
  distinct(all, cv) %>% 
  ggplot(aes(x=all, y=cv ))+
  geom_col(alpha=0.8, color="gray10")+
  scale_y_continuous(n.breaks=10)+
  coord_fixed(ratio = 15)+
  theme_bw()+ theme(text = element_text(size = 9, family="sans", color="black"), #stylisation
        axis.text = element_text(size=9, color="black"),
        axis.title=element_text(size=9, face = "bold"),
        plot.subtitle=element_text(size=10.5),
        plot.title=element_text(size=11, face = "bold"), 
        legend.position ="none") -> cv.all.1

df_psd.bulk.zfish_atac.lg1x %>% 
  group_by(celltype) %>% 
  distinct(cv_celltype, cv) %>% 
  ggplot(aes(x=celltype, y=cv_celltype, fill=celltype))+
  geom_col(alpha=0.8, color="gray10")+
  scale_y_continuous(n.breaks=10)+
  theme_bw()+ theme(text = element_text(size = 9, family="sans", color="black"), #stylisation
        axis.text = element_text(size=9, color="black"),
        axis.title=element_text(size=9, face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        plot.subtitle=element_text(size=10.5),
        plot.title=element_text(size=11, face = "bold"), 
        legend.position ="none") ->cv.celltype.1

cv.all.1|cv.celltype.1
```


## 3.5 Count Distribution by cell type

```{r}
df_psd.bulk.zfish_atac.lg1x %>% 
  mutate(all="all") %>% 
 ggplot( aes( x=all, y=pk_counts))+
  geom_violin(color="grey20", alpha=0.3, fill="gray30", trim=F)+
  coord_fixed(ratio = 0.5)+
  theme_bw()+ theme(text = element_text(size = 4, family="sans", color="black"), #stylisation
        axis.text = element_text(size=9, color="black"),
        axis.title=element_text(size=9, face = "bold"),
        plot.subtitle=element_text(size=10.5),
        plot.title=element_text(size=11, face = "bold"), 
        legend.position ="none") -> v.plot.all.1


 ggplot(df_psd.bulk.zfish_atac.lg1x, aes(x=celltype, y=pk_counts, fill = celltype))+
  geom_violin(color="grey20", alpha=0.3, trim=F)+
   coord_equal(ratio = 0.8)+
  theme_bw()+ theme(text = element_text(size = 4, family="sans", color="black"), #stylisation
        axis.text = element_text(size=9, color="black"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title=element_text(size=9, face = "bold"),
        plot.subtitle=element_text(size=10.5),
        plot.title=element_text(size=11, face = "bold"), 
        legend.position ="none") -> v.plot.cells.1
 
 v.plot.cells.1
```

side by side
```{r}
v.plot.all.1|v.plot.cells.1
```


The data is hardly visible so I made the following adjustments: boxplot and scale limits set to 0 and 0.1

The resulting plot can be seen down below. 
```{r}
df_psd.bulk.zfish_atac.lg1x %>% 
  mutate(all="all") %>% 
 ggplot( aes( x=all, y=pk_counts))+
  geom_violin(color="grey20", alpha=0.3, fill="gray30", trim=F)+
  geom_boxplot(width = 0.07)+
  stat_boxplot(geom ='errorbar', width=0.2, position=position_dodge(0.5))+
  scale_y_continuous(breaks = breaks_pretty(n=20))+
  coord_fixed(ratio = 0.8)+
  theme_bw()+ theme(text = element_text(size = 9, family="sans", color="black"), #stylisation
        axis.text = element_text(size=9, color="black"),
        axis.title=element_text(size=9, face = "bold"),
        plot.subtitle=element_text(size=10.5),
        plot.title=element_text(size=11, face = "bold"), 
        legend.position ="none") -> v.plot.all.box.1


 ggplot(df_psd.bulk.zfish_atac.lg1x, aes(x=celltype, y=pk_counts, fill = celltype))+
  geom_violin(color="grey20", alpha=0.3, trim=F)+
   geom_boxplot(width = 0.07)+
  stat_boxplot(geom ='errorbar', width=0.2, position=position_dodge(0.5))+
   scale_y_continuous(breaks = breaks_pretty(n=20))+
  # coord_equal(ratio = 20)+
   #facet_wrap(facets=vars(celltype), scales="free_x")+
  theme_bw()+ theme(text = element_text(size = 9, family="sans", color="black"), #stylisation
        axis.text = element_text(size=9, color="black"),
        #axis.text.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title=element_text(size=9, face = "bold"),
        plot.subtitle=element_text(size=10.5),
        plot.title=element_text(size=11, face = "bold"), 
        legend.position ="none") -> v.plot.cells.box.1
 
 v.plot.all.box.1|v.plot.cells.box.1
```

```{r}
v.plot.cells.box.1
```


